---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(realEstAnalytics)
set_zillow_web_service_id('X1-ZWz181enkd4cgb_82rpe')
YOURAPIKEYHERE = getOption('ZillowR-zws_id')
```
# realEstAnalytics

<!-- badges: start -->
[![Travis build status](https://travis-ci.org/xiyuansun/realEstAnalytics.svg?branch=master)](https://travis-ci.org/xiyuansun/realEstAnalytics)
[![Codecov test coverage](https://codecov.io/gh/xiyuansun/realEstAnalytics/branch/master/graph/badge.svg)](https://codecov.io/gh/xiyuansun/realEstAnalytics?branch=master)
<!-- badges: end -->

The goal of realEstAnalytics is to provide an R function for each zillow API service, making it easy to make API calls and process the responses into R-friendly data structures. 

## Website
The package is hosted at https://www.realestanalyticsr.com/

## Installation

You can install the released version of realEstAnalytics from [Github](https://github.com) with:

``` {r, eval=FALSE}
library(devtools)
devtools::install_github("xiyuansun/realEstAnalytics")
```

## Example

Here are some basic examples of interfacing with Zillow's API in `R`. All calls to the API require a unique Zillow API key, which you can acquire by signing up at https://www.zillow.com/howto/api/APIOverview.htm .

First, you should always set your API key, also known as a Zillow Web Service ID (ZWSID):

```{r, eval=FALSE}
library(realEstAnalytics)

#set the ZWS_ID
set_zillow_web_service_id('YOUR_API_KEY')
```
```{r}
#retrieve the current ZWS_ID in use
zapi_key = getOption('ZillowR-zws_id')
```

You can get basic information on a property based on its address using `GetDeepSearchResults`. The return is a dataframe with the property's estimated value as well as other characteristics (i.e., bedrooms, bathrooms, etc.). To get results for more than one address at once, we recommend `GetDeepSearchResults_dataframe` rather than using an `apply()` or other mapping function.
```{r}
GetDeepSearchResults('600 S. Quail Ct.', city='Newton',state='KS', zipcode=NULL,
                     api_key=getOption('ZillowR-zws_id'))
```

Use the `GetComps` or `GetDeepComps` to get comparable properties for a given Zillow Property ID (limit 25 comparables). The return is a data frame with just the comparable addresses and their Zestimate values, with more property information available with `GetDeepComps`.
```{r}
#retrieve the zpid from GetDeepSearchResults
zpidex <- GetDeepSearchResults('600 S. Quail Ct.', zipcode=67114,
                     rentzestimate=TRUE, api_key=getOption('ZillowR-zws_id'))$zpid

#GetComps for the '600 S. Quail Ct.' address
GetComps(zpidex, count=10, rentzestimate=TRUE, api_key = getOption('ZillowR-zws_id'))
```

You can get the Zestimate (Zillow's estimated home value) with `GetZestimate`. The return is a data frame with the Zillow-estimated value of the home, either the property value or the estimated rental value (if `rentzestimate=TRUE`). This function works with either a single Zillow property ID or a vector of IDs:
```{r}
GetZestimate(zpids= zpidex ,
             rentzestimate=TRUE , api_key=getOption('ZillowR-zws_id'))
```

There are also a few options for retrieving non-API data. Get the time series of property values aggregated by region and type with `get_rental_listings` and/or `get_ZHVI_series`. These return dataframes of time series data for the selected geographic aggregation level and property type, read from a static .csv hosted by Zillow. For options on the arguments, see the introduction vignette.
```{r}
#Pull the data by state and zipcode for 4 bedrooms
cityseries <- get_ZHVI_series(bedrooms=4,geography="Zip")

head(cityseries)
```

Please see the vignette (under articles) to see further documentation and examples of `realEstAnalytics`' capabilities.
